Unsupervised training of maximum-entropy models for lexical selection in rule-based machine translation

### 2015-02-09 2015-eamt-melsa/2015-eamt-melsa.tex
some reference culling
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58779 72bbbca6-d526-0410-a7d9-f06f51895060

Then \cite{berger1996} illustrated the use of maximum-entropy classifiers
on the specific problem of lexical selection in IBM-style word-based
statistical machine translation. Other authors
\citep{carbonell06,melero07a} have used target-language models to rank
the translations resulting from all possible combinations of lexical
selections.  Nowadays, in state-of-the-art phrase-based statistical
machine translation \citep{koehn2009statistical}, lexical selection is
taken care of by a combination of the translation model and the
language model. The translation model provides probabilities of
translation between words or word sequences (often referred to as
\emph{phrases}) in the source and target language. The target-language
model provides probabilities of word sequences in the target language.
As regards rule-based machine translation, \cite{sanchez07a} proposed 
to use information from the target language to train a statistical 
lexical-selection module but did not provide any experiments. 
\todo[inline]{FMT: integrate \citep{sanchez07a} somewhere here}
\cite{marechek10} trained a maximum-entropy lexical selector for their TectoMT rule-based system
using a bilingual corpus.  More recently, \cite{tyers12a} presented a
method of lexical selection for rule-based machine translation based
on rules which select or remove translations in fixed-length contexts,
along with a training method for learning the rules from a
word-aligned parallel corpus. The principal disadvantage with this
method is that it relies on a parallel corpus for the language pair,
something which is not always available.

@@

\subsubsection{Machine translation performance} 
\comment{This part of the text needs compressing and merging} This is
an extrinsic evaluation, which ideally would test how much the system
improves as regards an approximate measurement of final translation
quality in a real system. For this task, we use the widely-used {\sc
  bleu} metric (see Section~\ref{sec:eval-metrics}). This is not ideal
for evaluating the task of a lexical selection module as the
performance of the module will depend greatly on (a) the coverage of
the bilingual dictionaries of the RBMT system in question, and (b) the
number of reference translations. It is also worth noting that
successful lexical selections may not lead to successful translations
due to inadequate transfer of morphological features. For example, if
we translate the Spanish phrase \emph{La gente le dice que no venga}
`The people tell him not to come', where \emph{decir} can be
translated as `say' or `tell' in English, and our machine translation
system generates `The people \emph{tells} him not to come', it would
be counted the same as if it were `The people \emph{says} him not to
come', although the lexical selection made was more
adequate. Generation errors --- where a word is found in the bilingual
dictionary, but not in the target language morphological dictionary
--- may also lead to the same problem.  The {\sc bleu} metric \citep{papineni02} is
included only as it is commonly used to evaluate MT systems. Other
metrics with similar performance to {\sc bleu} such as {\sc meteor}
\citep{lavie09} and {\sc nist} \citep{doddington02} are not included
as we have chosen to focus on lexical-selection performance.

@@

\item \textbf{Target language model} (TLM). One method of lexical
  selection is to use the existing machine translation system to
  generate all the possible translations for an input sentence, and
  then score these translations \emph{on-line} on a model of the
  target language. The highest scoring sentence is then picked as
  output. This is the method used by \cite{carbonell06} and
  \cite{melero07a}.  We used a five-gram language model of surface
  forms generated from the target-language side of the parallel
  corpus. %\comment{FSM: Why not a much larger corpus? (I reviewer may ask)}
%(see section~\ref{sec:eval-corpus}). 
  This was the best performing system when we compared different
  approaches, but is impractical for real-world MT because the number
  of translations to perform grows exponentially with the length of
  the sentence.  It represents the best that can be achieved using the
  current systems without access to a parallel corpus. In our work, it
  was implemented with the IRST language modelling toolkit
  \cite{federico08a}.
\end{itemize}

>>>>

Then \cite{berger1996} illustrated the use of maximum-entropy classifiers
on the specific problem of lexical selection in IBM-style word-based
statistical machine translation. Other authors
\citep{melero07a} have used target-language models to rank %@@@ carbonell06
the translations resulting from all possible combinations of lexical
selections.  Nowadays, in state-of-the-art phrase-based statistical
machine translation \citep{koehn2009statistical}, lexical selection is
taken care of by a combination of the translation model and the
language model. The translation model provides probabilities of
translation between words or word sequences (often referred to as
\emph{phrases}) in the source and target language. The target-language
model provides probabilities of word sequences in the target language.
As regards rule-based machine translation, \cite{sanchez07a} proposed 
to use information from the target language to train a statistical 
lexical-selection module but did not provide any experiments. 
\todo[inline]{FMT: integrate \citep{sanchez07a} somewhere here}
\cite{marechek10} trained a maximum-entropy lexical selector for their TectoMT rule-based system
using a bilingual corpus.  More recently, \cite{tyers12a} presented a
method of lexical selection for rule-based machine translation based
on rules which select or remove translations in fixed-length contexts,
along with a training method for learning the rules from a
word-aligned parallel corpus. The principal disadvantage with this
method is that it relies on a parallel corpus for the language pair,
something which is not always available.

@@

\subsubsection{Machine translation performance} 
\comment{This part of the text needs compressing and merging} This is
an extrinsic evaluation, which ideally would test how much the system
improves as regards an approximate measurement of final translation
quality in a real system. For this task, we use the widely-used {\sc
  bleu} metric (see Section~\ref{sec:eval-metrics}). This is not ideal
for evaluating the task of a lexical selection module as the
performance of the module will depend greatly on (a) the coverage of
the bilingual dictionaries of the RBMT system in question, and (b) the
number of reference translations. It is also worth noting that
successful lexical selections may not lead to successful translations
due to inadequate transfer of morphological features. For example, if
we translate the Spanish phrase \emph{La gente le dice que no venga}
`The people tell him not to come', where \emph{decir} can be
translated as `say' or `tell' in English, and our machine translation
system generates `The people \emph{tells} him not to come', it would
be counted the same as if it were `The people \emph{says} him not to
come', although the lexical selection made was more
adequate. Generation errors --- where a word is found in the bilingual
dictionary, but not in the target language morphological dictionary
--- may also lead to the same problem.  The {\sc bleu} metric \citep{papineni02} is
included only as it is commonly used to evaluate MT systems.

%@@@ Other
%@@@ metrics with similar performance to {\sc bleu} such as {\sc meteor}
%@@@ \citep{lavie09} and {\sc nist} \citep{doddington02} are not included
%@@@ as we have chosen to focus on lexical-selection performance.

@@

\item \textbf{Target language model} (TLM). One method of lexical
  selection is to use the existing machine translation system to
  generate all the possible translations for an input sentence, and
  then score these translations \emph{on-line} on a model of the
  target language. The highest scoring sentence is then picked as
  output. This is the method used by %@@@ \cite{carbonell06} and
  \cite{melero07a}.  We used a five-gram language model of surface
  forms generated from the target-language side of the parallel
  corpus. %\comment{FSM: Why not a much larger corpus? (I reviewer may ask)}
%(see section~\ref{sec:eval-corpus}). 
  This was the best performing system when we compared different
  approaches, but is impractical for real-world MT because the number
  of translations to perform grows exponentially with the length of
  the sentence.  It represents the best that can be achieved using the
  current systems without access to a parallel corpus. In our work, it
  was implemented with the IRST language modelling toolkit
  \cite{federico08a}.
\end{itemize}
